---
title: Git
author: Alan Ryan
date: '2018-05-09'
slug: git
categories: []
tags: []
---
**Git is a version control system used to track changes in computer files and to coordinate work among multiple users.**  
**After installing Git you can initialize a repository:**  
```{r eval=FALSE, include=TRUE}
$ git init
```
**or alternatively, *clone* an existing repository:**  
```{r eval=FALSE, include=TRUE}
$ git clone https://github.com/al1973/piecemeal.git
```
**As an example, add three new files and then commit them to the repository:**  
```{r eval=FALSE, include=TRUE}
$ git add one two three
$ git commit -m 'First commit'
```
**You get a *master* branch pointed to by *HEAD*:**  
```{r echo=FALSE}
library(DiagrammeR)
grViz("
digraph {
  graph [overlap = true, fontsize = 12, rankdir = RL]
  node [shape = box, fontname = Helvetica, style = filled]
  node[fillcolor = yellow]  
  'one'; 'two'; 'three'
  node[shape = oval, fillcolor = orange]
  'HEAD'
  node[fillcolor = green]
  'master'
  edge[color = grey]
  'master'->'three' 'HEAD'->'master' 'three'->'two' 'two'->'one'
}
")
```
**Create a new branch called *testing*:** 
```{r eval=FALSE, include=TRUE}
$ git branch testing  
```
**Switch to the *testing* branch:** 
```{r eval=FALSE, include=TRUE}
$ git checkout testing
```
```{r echo=FALSE, fig.align='left'}
grViz("
digraph {
  graph [overlap = true, fontsize = 12, rankdir = RL]
  node [shape = box, fontname = Helvetica, style = filled]
  node[fillcolor = yellow]  
  'one'; 'two'; 'three'
  node[shape = oval, fillcolor = orange]
  'HEAD'
  node[fillcolor = green]
  'testing'; 'master'
  edge[color = grey]
  'master'->'three' 'testing'->'three' 'HEAD'->'testing' 'three'->'two' 'two'->'one'
}
")
```
  **Now *HEAD* points to the *testing* branch.**    
**Do another commit:**  
```{r echo=FALSE}
grViz("
digraph {
  graph [overlap = true, fontsize = 10, rankdir = RL]
  node [shape = box, fontname = Helvetica, style = filled]
  node[fillcolor = yellow] 
  'one'; 'two'; 'three'; 'four' 
  node[shape = oval, fillcolor = orange]
  'HEAD'
  node[fillcolor = green]
  'testing'; 'master'
  edge[color = grey]
  'master'->'three' 'testing'->'four' 'HEAD'->'testing' 'four' -> 'three' 'three'->'two'          'two'->'one'
}
")
```
**Switch back to the *master* branch:**  
```{r eval=FALSE, include=TRUE}
$ git checkout master
```
```{r echo=FALSE}
library(DiagrammeR)
grViz("
digraph {
  graph [overlap = true, fontsize = 10, rankdir = RL]
  node [shape = box, fontname = Helvetica, style = filled]
  node[fillcolor = yellow] 
  'one'; 'two'; 'three'; 'four' 
  node[shape = oval, fillcolor = orange]
  'HEAD'
  node[fillcolor = green]
  'testing'; 'master'
  edge[color = grey]
  'master'->'three' 'testing'->'four' 'HEAD'->'master' 'four'->'three' 'three'->'two'             'two'->'one'
}
")
```
**Make another change and commit:**  
```{r echo=FALSE}
grViz("
digraph {
  graph [overlap = true, fontsize = 10, rankdir = RL]
  node [shape = box, fontname = Helvetica, style = filled]
  node[fillcolor = yellow] 
  'one'; 'two'; 'three'; 'four'; 'five'
  node[shape = oval, fillcolor = orange]
  'HEAD'
  node[fillcolor = green]
  'testing'; 'master'
  edge[color = grey]
  'master'->'five' 'testing'->'four' 'HEAD'->'master' 'five'-> 'three' 'four' -> 'three'          'three'->'two' 'two'->'one'
}
")
```
**When you are ready you can merge the 2 branches together.**  
**Use the *merge* command to combine the *testing* branch with the *master* branch and then delete the *testing* branch:**  

```{r eval=FALSE, include=TRUE}
$ git merge testing
$ git branch -d testing
```
```{r echo=FALSE}
grViz("
digraph {
  graph [overlap = true, fontsize = 10, rankdir = RL]
  node [shape = box, fontname = Helvetica, style = filled]
  node[fillcolor = yellow] 
  'one'; 'two'; 'three'; 'four'; 'five'
  node[shape = oval, fillcolor = orange]
  'HEAD'
  node[fillcolor = green]
  'master'
  edge[color = grey]
  'master'->'five' 'five'->'four' 'HEAD'->'master' 'four'-> 'three' 'three'->'two' 'two'->'one'
}
")
```