---
title: Text Analysis - Moby Dick
author: Alan Ryan
date: '2018-02-20'
slug: text-analysis-moby-dick
categories: []
tags: []
---
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidytext)
library(gutenbergr)
library(tidyverse)
library(stringr)
library(wordcloud)
```
*Moby Dick* is Herman Melville's epic novel of 1851. It tells the story of Captain Ahab's pursuit of a great white sperm whale that has bitten off his leg. He risks his own life and that of his crew on the whaling ship *Pequod*. He is gripped by a narcissistic rage in his single-minded voyage of revenge. The tale is narrated by *Ishmael*, taking part in his first whaling expedition, and we encounter a multi-national crew including *Queequeg*, *Starbuck*, *Stubb*, *Tashtego*, *Flask* and *Daggoo*. The story is interspersed with detailed chapters on whales, almost in the form of a mini encyclopedia.  
Below is an analysis of the text:  

First, download the book from [Project Gutenberg](https://www.gutenberg.org/) and display the output:

```{r echo=FALSE, message=FALSE}
moby <- gutenberg_download(2489) %>%
        mutate(linenumber = row_number(),
               chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                       ignore_case = TRUE))))
select(moby, -gutenberg_id)
```

Then, tidy the text into a more manageable format

```{r echo=FALSE, message=FALSE}
tidy_moby <- moby  %>%
        unnest_tokens(word, text) %>% 
        select(-gutenberg_id)
knitr::kable(head(tidy_moby, n=10), align=c(rep('c', 5)))
```

Next, remove the stop words -- the uninteresting, common words, such as: I, me, my, myself, we, our, ours, ourselves, you... 

```{r echo=FALSE, message=FALSE}
data(stop_words)

tidy_moby <- tidy_moby %>%
        anti_join(stop_words)%>%
        count(word, sort = TRUE)
```

Of the remaining words, find the book's most frequent and display them in a word cloud:

```{r echo=FALSE, results='hide'}
with(tidy_moby, wordcloud(word, n, max.words = 100))
```
The R code used is available on [github](https://github.com/al1973/piecemeal/blob/master/content/post/2018-04-22-text-analysis.Rmd)